\documentclass[output=paper
,modfonts
,nonflat]{langsci/langscibook} 
%\bibliography{localbibliography}
%\documentclass[]{article}
%\input{localhyphenation.tex}
%\input{localpackages.tex}
%\input{localcommands.tex} 


\title{Long Distance Agreement and Locality: A Reprojection Approach}

\author{Kristin B\"orjesson\affiliation{Universit{\"a}t Leipzig}\and Gereon M\"uller\affiliation{Universit\"at Leipzig}} 

\abstract{Based on the assumption that all cases of long distance
agreement  should be analyzed as involving only local agreement,
and based on the observation that all existing approaches to long-distance agreement in
terms of local agreement face substantial empirical and/or conceptual
problems, the present paper sets out to develop a radically new
approach to long-distance agreement. We suggest that long-distance agreement can and should be analyzed as a
strictly local operation taking place early in the derivation, and
giving rise to a counter-bleeding effect (i.e., apparent non-locality)
later in the derivation as a consequence of regular syntactic
structure building. More specifically, we argue that long-distance agreement involves (a)
(what will become) a matrix verb V$\_$1 which enters the syntactic
derivation as part of a complex predicate V$\_$1-V$\_$2 that is merged with
the embedded internal argument, agreeing with it locally early in the
derivation; and (b) subsequent reprojection movement of V$\_$1 out of
V$\_$2's clause, which eventually produces a biclausal structure (and
thereby leads to a counter-bleeding effect with agreement). Empirical
evidence for the new approach mainly comes from Nakh-Daghestanian
languages, among them Hinuq, Khwarshi, and Tsakhur.
}

\begin{document}

\maketitle

\newcommand{\refprefix}{ex:mueller:}
\section{Introduction}

Long-distance agreement is a phenomenon where agreement seems to take
place in a non-local configuration, that is, across a clause boundary.
More specifically, in cases of long-distance agreement, the verb in
the matrix sentence agrees with respect to $\phi$-features with an
argument of the verb in an embedded sentence. A language that show
this phenomenon is Hindi-Urdu (see \cite{Mahajan:90},
\cite{Butt:95,Butt:08}, \cite{Bhatt:05}, and \cite{Chandra:05}, among
others). A relevant pair of examples is given in (\Next). In Hindi, a DP
qualifies as an agreement controller if it is not overtly case-marked (if
both an external and an internal argument fail to be overtly
case-marked, agreement is with the subject). Long-distance agreement
is optional here: Either there is agreement of both the matrix verb
and the embedded verb with the embedded absolutive object DP (as in
\Next[b]), or the verbs show default agreement (as in \Next[a]).

\ea\label{ex:mueller:1}
\ea 
\gll  Raam-ne [$\alpha$ rotii khaanaa~] chaahaa \\ 
      Ram.{\scshape masc}.-{\scshape erg} {} bread.{\scshape fem} eat.{\scshape inf}.{\scshape masc} want.{\scshape perf.pst.masc} \\ 
\glt  `Ram wanted to eat bread.'
\ex 
\gll Raam-ne [$\alpha$ rotii khaanii~] caahii \\ 
     Ram.{\scshape masc}-{\scshape erg} {} bread.{\scshape fem} eat.{\scshape inf}.{\scshape fem} want.{\scshape perf.pst.fem} \\ 
\glt `Ram wanted to eat bread.'
\z
\z
Long-distance agreement is also widespread in Nakh-Daghestanian
languages. A relevant pair of examples from Tsez is given in (\Next)
(see \cite{Polinsky&Potsdam:01}). Agreement with respect to gender
(III, in the case hand) is controlled by
absolutive DPs in Tsez. It always shows up on the embedded verb (if
that verb can host overt agreement morphology in principle), and may
then optionally also show up on the matrix verb (as in \Next[b]);
alternatively, there is no long-distance agreement, and the matrix
verb exhibits default (IV) agreement marking (as in \Next[a]).\footnote{As
  noted by \cite{Bhatt:05}, some varieties of Hindi behave similarly
  in that they also exhibit an asymmetry between agreement with matrix
  as opposed to embedded verbs (such that embedded verbs can agree
  while matrix verbs do not have to), whereas other varieties of Hindi show a
  strict one-to-one correspondence (such that embedded verb agreement
  implies matrix verb agreement), as presupposed in the main text above.}

\ea\label{ex:mueller:2}
\ea
\gll     Eni-r [$\alpha$ u\v{z}\={a} magalu b-\={a}c'-ru-\l i~] r-iy-xo \\
         mother-{\scshape dat} {} boy-{\scshape erg} bread.{\scshape iii.abs} {\scshape iii}-eat-{\scshape pstprt-nmlz} {\scshape iv}-know-{\scshape prs} \\
\glt     `The mother knows that the boy ate the bread.'
\ex
\gll     Eni-r [$\alpha$ u\v{z}\={a} \label{2-b}magalu b-\={a}c'-ru-\l i~] b-iy-xo \\
         mother-{\scshape dat} {} boy-{\scshape erg} bread.{\scshape iii.abs} {\scshape iii}-eat-{\scshape pstprt-nmlz} {\scshape iii}-know-{\scshape prs} \\
\glt      `The mother knows that the boy ate the bread.'
\z
\z
Another example for long-distance agreement in Nakh-Daghestanian
languages comes from Hinuq; see (\Next[a--b]) (from
\cite{Forker:11}). Again, gender agreement is controlled by an
absolutive DP, and gender-based long-distance agreement is optional.

\ea\label{ex:mueller:3}
\ea
\gll Sa\textbarglotstop ida-z r-eq'i-yo [$\alpha$ Madina-y \textgamma i ga:-s-\textbeltl i~]\textsubscript{V} \\
     Saida-{\scshape dat} {V}-know.{\scshape prs} {} Madina-{\scshape erg} milk(IV).{\scshape abs} drink-{\scshape res-abst} \\ 
\glt `Saida knows that Madina drank milk.'
\ex
\gll  Sa\textbarglotstop ida-z y-eq'i-yo [$\alpha$ \label{3-b}Madina-y \textgamma i ga:-s-\textbeltl i~]\textsubscript{V} \\
      Saida-{\scshape dat} {IV}-know.{\scshape prs} {} Madina-{\scshape erg} milk({IV}).{\scshape abs} drink-{\scshape res-abst} \\
\glt  `Saida knows that Madina drank milk.'
\z
\z
Other languages exhibiting long-distance agreement are Itelmen (see
\cite{Bobaljik&Wurmbrand:05}; a relevant pair of examples is given in
\Next), Innu-aim\^{u}n (see \cite{Branigan&MacKenzie:02}, with
relevant examples in \NNext), Passamaquoddy (see \cite{Bruening:01}),
Chukchee (see \cite{Boskovic:07}) and Blackfoot (see \cite{Bliss:09}).

\ea\label{ex:mueller:4}
\ea 
\gll Na netxa-in [$\alpha$ kma je\textipa{B}na-s~] \\
     he forget-{{\scshape 3sg.subj(intrans)}} {} me meet-{\scshape inf} \\
\glt  `He forgot to meet me.'
\ex
\gll Na \textipa{@}ntxa-\textipa{B}um=nm [$\alpha$ kma je\textipa{B}na-s~] \\
     he forget-{{\scshape 1sg.obj=3.cl}} {} {me} meet-{\scshape inf} \\
\glt `He forgot to meet me.'
\z
\z

\ea\label{ex:mueller:5}
\ea 
\gll Ni-tshissenit-en [$\alpha$ P\^{u}n k\^{a}-m\^{u}pisht-\^{a}shk~] \\
     1-know-{{\scshape ti}} {} Paul {\scshape prt}-visited-2/{\scshape inv} \\
\glt `I know that Paul visited you.'
\ex
\gll Ni-tshissenim-{\^a}u [$\alpha$ P\^{u}n k\^{a}-m\^{u}pisht-\^{a}shk~] \\
     1-know-{3} {} {Paul} {\scshape prt}-visited-2/{\scshape inv} \\
\glt `I know that Paul visited you.'
\z
\z
While long-distance agreement is optional in all these languages,
there are also some languages where long-distance agreement is
obligatory in certain environments; this holds, e.g., for Icelandic,
Kutchi Gujarati and Chamorro.  In this paper, we will concentrate on
those languages in which long-distance agreement seems
optional. Still, as with many other cases of syntactic optionality, it
turns out that in all these cases, the choice of long-distance
agreement in a sentence goes along with an interpretation of the
controller of long-distance agreement as having a particular
information structural status, i.e., an interpretation that the other
member of the sentence pair -- that with (only) local agreement --
lacks. 

The central challenge posed by long-distance agreement for syntactic
theory is, of course, the apparent non-locality of the operation. More
specifically, the embedded DP that controls the agreement would seem
to be separated by a clause-like constituent ($\alpha$ in the examples
above) from the matrix verb, and therefore be too far away to permit
establishing a local relation.  This is potentially problematic
because most current syntactic theories do indeed postulate that
syntactic operations (like agreement) are highly local (e.g., this
holds for the Minimalist Program, HPSG, Categorial Grammar, and Optimality
Theory); in line with this, in all these theories, apparently
non-local dependencies like long-distance movement, long-distance
reflexivization, long-distance case assignment, sequence of tense, and
switch reference have successfully been reanalyzed as involving only
fairly local operations (see \cite*{Alexiadouetal:12} for an
overview).

Against the background of the Minimalist Program, the question raised
by long-distance agreement is how it can be ensured that the basic
locality requirement imposed by the Phase Impenetrability Condition
(PIC; \cite{Chomsky:00,Chomsky:01,Chomsky:08,Chomsky:13}) in (\Next) is
respected by the operation. 

\ea\label{ex:mueller:6} {\itshape Phase Impenetrability Condition} \label{pic1}(PIC):\\
  In phase $\alpha$ with head H, the domain of H is not accessible to
  operations outside $\alpha$; only H and its edge are accessible to such
  operations.
\z


If $\alpha$ in \ref{1}--\ref{5} qualifies as a phase in the sense of
(\Last), the very existence of long-distance agreement seems to be at
variance with the PIC.\footnote{\label{fn1}As a matter of fact, simple cases of T
  agreeing with a nominative object DP in VP in a language like
  Icelandic are already problematic from the point of view of the PIC
  in \ref{pic1}. For this reason, \cite{Chomsky:01} also envisages a
  second, somewhat more liberal version of the PIC, where a phase
  domain becomes opaque only when the next phase is reached. However, even
  this less restrictive version of the PIC would not suffice to
  straightforwardly derive the possibility of long-distance
  agreement. Furthermore, one might argue that the head actually
  responsible for this agreement into VP is not T but v. We will
  abstract away from this issue in what follows, and presuppose the
  PIC in \ref{pic1} for the remainder of the paper.}

In what follows, we will argue that the PIC-induced locality problem
with long-distance agreement is real, and has not been
convincingly solved yet in any of the approaches to long-distance
agreement that have been developed so far. In view of this state of
affairs, we will propose a radically new approach in terms of complex
predicate formation plus reprojection where long-distance agreement
can be analyzed as a strictly local operation: The new analysis  involves (i) a verb
V$\_$1 which enters the syntactic derivation as part of a complex
predicate V$\_$1-V$\_$2 that is merged with the embedded internal argument,
agreeing with it locally early in the derivation; and (ii) subsequent
reprojection movement of V$\_$1 out of V$\_$2's clause, which eventually
produces a biclausal structure (and thereby leads to a
counter-bleeding effect with agreement). 

The article is structured as follows. In section 2, we will sketch the
different types of existing approaches to the phenomenon of
long-distance agreement. Given the PIC-based premise of strictly local
application of all syntactic operations, we will point out individual
problems the different approaches face as well as introduce some new
data that turns out to be problematic for almost all of them. We will
then go on to introduce the new approach in terms of reprojection in
section 3. In section 4, which concludes the article, we will
summarize the main features of the new approach and provide an outlook
into how it might also be put to use in other contexts involving
extraction from DPs where there is evidence for an extremely local
relation of two heads that show up in two separate domains in
syntactic output structures.
  

\section{Existing Analyses}

At least for our present purposes, four different kinds of analysis
of long-distance agreement can be distinguished. We will refer to
these as (i) non-local analyses (where long-distance agreeement can
apply in a non-local fashion),  (ii) small structure analyses (where
there is no phase boundary), (iii) cyclic Agree analyses (where the
information relevant for agreement is locally passed on through the
tree, originating with the controller and ultimately reaching the
matrix verb), and (iv) feeding analyses (where movement
of the controller makes 
local agreement possible). We will discuss these four analysis types in
turn.\footnote{We will not consider a fifth type of analysis, where
  the matrix verb locally agrees with a covert pronoun, which in turn
  is coindexed (and therefore shares $\phi$-features) with an
  embedded DP. As shown in \cite{Polinsky&Potsdam:01} and
  \cite{Bhatt&Keine:16:lon}, such `proxy agreement' is not a viable
  alternative in general. (Also, it is worth noting that such an
  analysis solves one locality problem (seemingly non-local agreement)
by shifting it to another, well-known locality problem (seemingly
non-local binding chains).)}

\subsection{Type (i): Non-Local Analyses}

Non-local analyses either assume that the locality constraint on Agree
is weaker than the original PIC (see \cite{Chomsky:01} and footnote
\ref{fn1}), or that Agree is not, in fact, subject to such a strict
locality constraint in the first place (see \cite{Sells:06} \cite{Boskovic:07},
\cite{Keine:16}).

For example, Bo\v{s}kovi{\'c}'s (2007) analysis relies on a revised
Agree operation, one which is not subject to either the  Activity
  Condition (\ref{AC} in the original Agree definition in
\ref{Agree}) or the PIC. 

\ea\label{ex:mueller:7} \label{Agree}{\itshape Agree} \\ $\alpha$ may Agree with $\beta$ iff
(a)--(d) hold.
\ea  $\alpha$ carries at least one unvalued and uninterpretable
      feature and $\beta$ carries a matching interpretable and valued
      feature 
\ex  $\alpha$ c-commands $\beta$
    \b. \label{AgreeClosest}$\beta$ is the closest goal to $\alpha$
\ex \label{AC}$\beta$ bears an unvalued uninterpretable feature.
\z
\z
Thus, the matrix verb can look all the way
down to the embedded absolutive DP to check its
$\phi$-feature. \citet{Boskovic:07} assumes that finite complement
clauses can in principle be CPs or TPs. Thus, in complement clauses in
which there is no evidence for a CP layer, one might as well asume
that those clauses actually lack it.
The idea, then, is that CPs block long-distance agreement in Tsez while TPs allow it. This
is because, Bo\v{s}kovi{\'c} (2007) claims, CPs, in contrast to TPs
may carry $\phi$-features that need to get checked. In case of Tsez
CPs (and those of the languages that lack long-distance agreement altogheter), he assumes
that they do carry such (default-valued) $\phi$-features, which makes it possible for
the matrix verb to locally agree with the CP as such, leading to local
agreement. Long-distance agreement in cases that involve CPs is impossible due to the
condition on Agree to involve that potential goal closest to the probe
(see \ref{AgreeClosest}).

Whatever the merits of this proposal, it is clear that it is
incompatible with the strict PIC in \ref{pic1}: 
 Since the potential long-distance agreement controller  in
transitive sentences is the internal argument bearing absolutive case, which is base-generated
in VP, the PIC will predict that long-distance agreement should not be
possible, independently of whether the embedded vP phase has both a TP
and a CP projection on top of it, or just a TP projection. 

\subsection{Type (ii): Small Structure Analyses}

In small structure analyses, it is argued that the configuration of matrix verb
and long-distance agree\-ment-trigger is local after all. The structure
of the complement clause is assumed to be smaller than might be
thought from the surface data (e.g., a VP in \cite{Boeckx:04}; an
InflP in \cite{Bhatt:05}).

Like \citegen{Boskovic:07} approach, \citegen{Bhatt:05} analysis is
    based on a revised Agree operation -- in this case, it is one which also is not
 subject to the {\itshape Activity Condition}, but which does respect the PIC.
To account for apparent long-distance agreement in Hindi, \citet{Bhatt:05} assumes
   that complement clauses to  matrix verbs that allow long-distance agreement are in fact
   only InflPs/VPs, which lack an external argument (i.e., they have
   no PRO) and are thus not phases. In contrast, long-distance
   agreement out of finite clauses is not possible in Hindi because
   finite complement clauses are CPs; the same prediction arises
   for infinitival structures that contain a PRO subject. 
In line with this, as far as the optionality of long-distance agreement in Hindi is
concerned, 
\citet{Bhatt:05} explains it by
   assuming that matrix verbs that allow long-distance agreement have an option of selecting
   either a restructuring infinitive or a non-restructuring infinitive, where the latter involves a syntatically projected PRO subject.
    This PRO intervenes between the matrix verb and the embedded
    object and, thus, blocks long-distance agreement (in the same way,
    a problem with the PIC would arise in this context). 

\citegen{Bhatt:05} analysis may work well for a language like Hindi,
but it does not carry over to long-distance agreement in languages of
the Nakh-Daghestanian type, where the external argument is clearly
present (bearing ergative) in the embedded clause; see, e.g.,
\ref{2-b} from Tsez and \ref{3-b} from Hinuq. Given the strict version
of the PIC in \ref{pic1}, it is clear that the presence of an external
argument DP uncontroversially implies the presence of a vP phase, and
this should make agreement of a matrix V with an internal argument DP
included in the complement domain of v impossible, independently of
whether vP qualifies as $\alpha$ (in the above sense) or not (i.e.,
independently of whether there is additional structure on top of vP in
the complement of the matrix V.\footnote{As a matter of fact, to
  account for evidence of the Nakh-Daghestanian type, 
\citet[791]{Bhatt:05} ultimately concludes that a feeding account of
along the lines of \cite{Polinsky&Potsdam:01}  is independently called
for; see below.}

\subsection{Type (iii): Cyclic Agree Analyses}

In cyclic Agree analyses, it is assumed that what looks like long-distance
agreement actually is to be decomposed into a series of shorter
agreement steps, all of of which obey strict locality. On this view,
first the embedded verb agrees with the embedded agreement controller
DP; second, the matrix verb agrees with the embedded verb; third, by
transitivity, this implies that the matrix verb will eventually agree
with the embedded DP, albeit indirectly. This kind of analysis has
been pursued by \cite{Butt:95}, \cite{Legate:05:pha}, \cite{Keine:08},
\cite{Preminger:09}, and \cite{Lahne:12}, among others. As an
illustration, consider the specific approach developed in
\cite{Legate:05:pha}.

The basic premise of this approach is that at no stage of the
derivation is there an Agree relation between the matrix verb and the
embedded DP.  Rather, the agreement controller DP's $\phi$-features
first valuate an $[u\phi]$ probe feature of a phase head, which by
definition (cf. the PIC in (\ref{pic1})) is also part of the higher
phase.  The matrix verb then probes the embedded phase head's
$\phi$-features.  Thus, the embedded phase head acts as a hinge
between the matrix and embedded domains.  This accounts for the
observation that long-distance agreement presupposes the existence of
local agreement in the embedded clause. 

The cyclic Agree approach solves the locality problem with
long-distance agreement in a very simple manner that directly
corresponds to the analogous (and by now well-established) treatment
of long-distance movement in terms of successions of smaller movement
steps.\footnote{As a matter of fact, an alternative local analysis of
  long-distance agreement that mimicks {\scshape slash} feature percolation
  as it has been proposed for movement dependencies (see
  \cite{Gazdar:81}) might in principle also be an option. This would
  then express a similarity of the two operations in long-distance
  contexts (viz., movement and agreement) even more
  straightforwardly. However, to the best of our knowledge, such an
  analysis has not yet been proposed. That notwithstanding, it would
  be subject to the same empirical problem mentioned in the main text
  below.}  However, there are both conceptual and empirical problems
raised by cyclic Agree approaches to long-distance agreement. On the
one hand, cyclic Agree is conceptually problematic from a minimalist
perspective, given standard assumptions about probe features, goal
features, and the Agree operation: It looks as though one and the same
set of $\phi$-features (on the phase head in the middle) must act as a
probe in one case, and as a goal in another (see \cite{Bhatt:05}). On
the other hand, there is an empirical problem (see
\cite{Polinsky&Potsdam:01}, \cite{Bhatt&Keine:16:lon}) that is due to the
fact that cyclic Agree approaches rely on transitivity. The problem is
that if two verbs V$\_$1 and V$\_$2 can in principle participate in local
agreement in some long-distance agreement constructions, and V$\_$2 and
DP$\_${abs} can participate in local agreement, then long-distance
agreement involving V$\_$1 and DP$\_${abs} must also be possible. However,
this is not always the  case: For instance,  in Tsez, long-distance
agreement, unlike local agremeent, requires DP$\_${abs} to
be a topic (see \cite{Polinsky&Potsdam:01}). This is shown by the
examples in (\Next), where (i) local agreement of the embedded predicate and
the absolutive DP as the agreement controller is possible throughout,
(ii) the matrix and embedded predicates can participate in
long-distance agremeent in principle, but (iii) long-distance
agreement in this configuration is blocked nevertheless because the
absolutive DP is not interpreted as a topic: It is interpreted as a
focus in (\Next[a]), and it shows up as an -- inherently
non-topicalizable -- reflexive pronoun in (\Next[c]). In both cases, the
matrix verb can only carry out agreement with the embedded clause
($\alpha$) itself; cf. (\Next[b--d]). 

\ea\label{ex:mueller:8}
\ea[*]{
 \gll   Eni-r [$\_${$\alpha$}$\_${\rm IV} t$'$ek-kin y-igu y\={a}\l -ru-\l i~] y-iy-xo\\
        mother-{\scshape dat} {} bookII.{\scshape abs}-{\scshape foc} II-good be-{\scshape pstprt-nmlz} II-know-{\scshape pres}\\
 \glt   `The mother knows that the {\scshape book} \rm is good.'}
\ex[]{ 
 \gll Eni-r [$\_${$\alpha$}$\_${\rm IV} t$'$ek-kin y-igu y\={a}\l -ru-\l i~] r-iy-xo\\
       mother-{\scshape dat} {} bookII.{\scshape abs}-{\scshape foc} II-good be-{\scshape  pstprt-nmlz} IV-know-{\scshape pres}\\
 \glt `The mother knows that the {\scshape book} \rm is good.'}
\ex[*]{
 \gll Eni-r [$\_${$\alpha$}$\_${\rm IV} u\v{z}-\={a} {nes\={a} \v{z}e}  \v{z}\={a}k$'$-ru-\l i~] {\O}-iy-xo \\
  mother-{\scshape dat} {} boy-{\scshape erg} {\scshape refl}.I.{\scshape abs} beat-{\scshape  pstprt-nmlz} I-know-{\scshape pres}\\
 \glt `The mother knows that the boy beat himself up.'}
\ex[]{
  \gll Eni-r [$\_${$\alpha$}$\_${\rm IV} u\v{z}-\={a} {nes\={a} \v{z}e} \v{z}\={a}k$'$-ru-\l i~] r-iy-xo \\
mother-{\scshape dat} {} boy-{\scshape erg} {\scshape refl}.I.{\scshape abs} beat-{\scshape pstprt-nmlz} IV-know-{\scshape pres}\\
 \glt `The mother knows that the boy beat himself up.'}
 \z
 \z

\subsection{Type (iv): Feeding Analyses}

\subsubsection{Movement Feeds Agreement}

Fourth and finally, \cite{Polinsky&Potsdam:01} have argued that a
local approach to long-distance agreement in Tsez is both technically
feasible and empirically supported according to which long-distance
agreement involves feeding of local agreement by movement. The basic
assumption is that the agreement controller (an absolutive DP in Tsez)
moves to a position in which it can locally agree with the matrix
verb. However, there are two complications to this simple
picture. First, Polinsky \& Potsdam present strong arguments against
the assumption that displacement of the agreement controller DP ends
up in the matrix clause itself. For one thing, all established
movement operations in Tsez are strictly clause-bound, so the 
operation that feeds long-distance agreement would be the only type of
movement that could leave a clause. For another,
\cite{Polinsky&Potsdam:01} observe that long-distance
agreement in Tsez never co-occurs with scope reversal; in other words,
a DP that participates in long-distance agreement with a matrix verb
can never take scope over quantifed items in the matrix clause.
This latter property is illustrated in (\Next): Independently of whether
long-distance agreement takes place (see \Next[b]) or not (see
\Next[a]), the embedded absolutive DP (with a universal quantifier in
\Next) cannot take scope over a  matrix subject (with an existential
quantifier in the case at hand). 

\ea \label{ex:mueller:9}  \label{scope32}
\ea
\gll Sis u\v{c}iteler [$\_${$\alpha$} \v{s}ibaw u\v{z}i \o -ik'ixosi-\textbeltl i] r-iy-xo \\
     one teacher {} every boy{\scshape i}.{\scshape abs} {\scshape i}-go-{\scshape nmlz} {\scshape iv}-know-{\scshape prs}  \\
\glt `Some teacher is such that he knows that every boy is going.'\\
     \char42`Every boy is such that some teacher knows that he is going.'
\ex
\gll Sis u\v{c}iteler [$\_${$\alpha$} \v{s}ibaw u\v{z}i \o -ik'ixosi-\textbeltl i] \o -iy-xo \\
     one teacher {} every boy{\scshape i}.{\scshape abs} {\scshape i}-go-{\scshape nmlz} {\scshape i}-know-{\scshape prs}  \\
\glt      `Some teacher is such that he knows that every boy is going.'\\
    \char42`Every boy is such that some teacher knows that he is going.'
\z
\z
From these considerations it follows that the postulated movement operation
cannot actually end up in the matrix clause in cases of long-distance
agreement. The second complication involves the overt/covert
distinction of movement operations. Since the absolutive DP that participates in
  long-distance agreement does not have to be overtly
  displaced and typically shows up in its in situ position (or, more generally, given
  that Tsez exhibits variable word order: it shows up in its unmarked position),
  it is clear that the movement operation that feeds long-distance
  agreement must be a covert one. 

Against this background, \citegen{Polinsky&Potsdam:01} proposal is
the following: There is  covert, information structure-driven
  movement of the long-distance agreement-controller into a higher domain (phase) of the
  same clause, and the position thus reached provides a  local enough
  configuration with the matrix verb to make  Agree with it possible.
More specifically, \citegen{Polinsky&Potsdam:01} analysis works as
follows. 

A crucial basic assumption is that the size of embedded clauses in
Tsez is variable. (\Next) gives the maximal syntactic structure for a
clause in Tsez. This structure is only fully  built up when needed;
i.e., clauses are CPs if they exhibit material that belongs in this
layer (e.g., a C head) but not otherwise; a TopP is projected if the
clause contains a topic; and so forth. 

\ea\label{ex:mueller:10} {\itshape Clause \label{Tsezstructure}structure for Tsez}:\\
%\begin{tabular}{ccccccccc} 
%& \node{0}{CP}\\[4mm]
%\node{1}{~\phantom{asdf }} & & \node{2}{C$'$}\\[4mm]
%&\node{3}{C} && \node{4}{TopP}\\[4mm]
%&&\node{5}{~\phantom{asdf }} & & \node{6}{Top$'$}\\[4mm]
%&&&\node{7}{Top} & & \node{8}{TP}\\[4mm]
%&&&&&\node{9}{DP$\_${erg} DP$\_${abs} V}\\
%\end{tabular}
%\nodeconnect{0}{1}\nodeconnect{0}{2}\nodeconnect{2}{3}\nodeconnect{2}{4}\nodeconnect{4}{5}
%\nodeconnect{4}{6}\nodeconnect{6}{7}\nodeconnect{6}{8}\nodetriangle{8}{9}
\z
On this  basis,  \cite{Polinsky&Potsdam:01} assume that, generally, long-distance agreement-allowing matrix
  verbs can select a TP as complement. 
However,  to derive long-distance agreement in Tsez, it is postulated
that a   topic-marked long-distance agreement-controlling element is covertly moved to the
  specifier of TopP in the left periphery of the complement clause. 
 This movement brings the triggering element into a sufficiently
  local relation to the matrix verb to allow the latter to check its
  uninterpretable $\phi$-features against those of the covertly
  topicalised element, resulting in long-distance agreement.
 Long-distance agreement is, thus, taken to be a reflex of the topic-status of the
  triggering element. 

 A CP is predicted to block long-distance agreement, as is a bare TP,
 given that a DP$\_${abs} agreement controller does not occupy SpecT (at
 this point the analysis is not fully PIC-compatible). The relation
 that \cite{Polinsky&Potsdam:01} assume to underlie long-distance
 agreement is  Head Government (not Agree, as in \ref{Agree}),
 which, following \cite{Rizzi:90}, is understood as in (\Next).

\ea\label{ex:mueller:11} {\itshape Head Government}:\\
X head-governs Y iff (a)--(d) hold:
\ea \label{list1}X $\in$ $\{$A, N, P, V, H[+{\rm tense}]$\}$.
\ex X m-commands Y.
\ex No barrier intervenes.
\ex Relativized Minimality is respected. 
\z
\z
It is then postulated that it can be derived from (\Last) that ``a head
governs its specifier, its complement, an element adjoined to its
complement, and the specifier of its complement''
\citep[627]{Polinsky&Potsdam:01}, but not, say, the specifier of
the complement of the head's complement. However, as a matter of fact,
  it is not quite clear why this should be the case. Consider (\Next).

\ea\label{ex:mueller:12} [XP X [ZP \label{vcp}Z [WP YP [W$'$ ...~]]]] \z
Does X head-govern YP in (\Last)? First, suppose that X is one of the
possible items mentioned in \ref{list1}. Second, X clearly m-commands
YP. Third, Relativized Minimality is respected in (\Last) because there
is no intervening phrase that could induce a Relativized Minimality
effect for X and YP (there is no phrase that c-commands YP and is
c-commanded by X -- ZP and WP both dominate YP). That leaves the
presence of a barrier as the only possible source of a failure of
head government of YP by X. Whether ZP or WP is a barrier in (\Last)
depends on the exact definition of this concept (which Polinsky and
Potsdam do not provide). The first thing to note is that both ZP and
WP are complements, i.e., sisters of X$^0$ categories. This will
suffice to exempt them from barrier status in most of the available
conceptions of barriers (see, e.g., \cite{Cinque:90}). In contrast,
according to the more complex, two-stage definition of barrier in
\cite{Chomsky:86}, ZP might in fact emerge as a barrier in (\Last) if WP
can be classified as a blocking category that passes on its status as
a ``virtual barrier'' to the phrase immediately above it. So it seems
that only under this complex approach, based on blocking categories
vs. real barriers, can it be derived that X does not head-govern YP in
(\Last). 

Based on the assumption that all agreement relations are subject to
(\LLast) (or at least the general consequences that (\LLast) is supposed
to have), a number of restrictions that Polinsky and Potsdam observe for
long-distance agreement in Tsez follow. First, a CP can never be
projected in long-distance agreement contexts because ``it would block
government of SpecTop by the verb''
\citep[638]{Polinsky&Potsdam:01}. Note  that this presupposes that
ZP (= CP) would indeed qualify as a barrier in (\Last) that makes
head government of YP (= DP$\_${abs} in SpecTop) by matrix V (= X)
impossible. (As we have just seen, this consequence is far from
straightforward.) However, with this qualification, it can be derived
that long-distance agreement is impossible (i) in the presence of a
wh-phrase in a clause (which inherently activates the CP layer,
whether or not wh-movement takes place overtly), and (ii) in the
presence of the element {\itshape \textcrlambda in}, which is assumed to be
a designated C element.\footnote{\label{whabs}As noted by
  \citet[fn. 20]{Polinsky&Potsdam:01}, long-distance agreement in the
  presence of a 
  wh-phrase would ceteris paribus be expected to be possible if the
  wh-phrase is itself the absolutive argument and occupies SpecC.}
 A third prediction is that long-distance
agreement with DP$\_${abs} is impossible if some other XP functions as
the topic in a clause.\footnote{Again, though, details of the account
  are somewhat unclear. One might think that this effect is due to
  intervention, i.e., the Relativized Minimality part of the
  definition of Head Government. However,  if there can only be one topic per clause in
  Tsez, and that is not DP$\_${abs}, then DP$\_${abs} can never reach the
  position (viz., SpecTop) it needs to reach to enable long-distance
  agreement, and a resort to intervention is not necessary. If, on the other hand, there can be more than one topic
  per clause, then it is not obvious why DP$\_${abs} should not qualify
  as the structurally highest one of them, thereby circumventing an
  intervention effect.}

These qualifications notwithstanding, \citegen{Polinsky&Potsdam:01}
analysis would seem to derive licit and illicit cases of long-distance
agreement in a very simple and elegant way that furthermore respects
locality considerations. Still, as shown in the following subsection,
there are conceptual and empirical problems with this approach.


\subsubsection{Problems with the Feeding Approach}

\subsubsubsection{The Nature of Covert Topic Movement}

First, as noted by \cite{Boskovic:07}, the crucial postulation of a
covert topicalisation operation for Tsez is far from innocuous. There
is virtually no independent evidence that such an operation
exists. Also, there is a real danger of an ordering paradox: It is not
really clear how {\itshape covert} movement at LF can trigger {\itshape overt}
agreement -- if the movement takes place at LF, it comes too
late.\footnote{This problem  can in principle be solved by assuming
  that covert movement is actually movement taking place in the
  narrow syntax, with the only difference to overt movement being that
  the lowest copy of a complex chain is subject to phonological
  realization (rather than the highest member, as with overt movement).
However, even if one were to adopt such an approach based on the copy
theory of movement, the intended effect does not seem to arise
anywhere else. As far as we can tell, other instances of covert
movement that have been suggested in the literature (e.g., for certain
cases of wh-in situ) never feed Agree operations. Thus, compare (i-a),
where overt wh-movement gives rise to new options for reflexivization
(assumed here to be an instance of Agree, see \cite{Reuland:11} for
discussion), with (i-b), where covert wh-movement fails to produce the
same effect (see \cite{Barss:86}). 

\ea\label{ex:mueller:Fn8}
\ea[]{John$\_$1 wonders [CP [DP which book about himself$\_$1~] Bill bought~]}
\ex[*]{John$\_$1 wonders [CP why Bill bought [DP a book about himself$\_$1~]]}
\z
\z
In the same way, covert wh-movement (unlike overt wh-movement) never seems to feed case
assignment or agreement in the world's languages. -- 
All  that said, \citet[626]{Polinsky&Potsdam:01} would seem to 
exclude a reinterpretation of covert movement as overt movement plus
pronunciation of the lower copy when they explicitly state that the
``syntactic agreement configuration between the probe and the
absolutive trigger is created at LF''.}


\subsubsubsection{Complementizers}

There are two comp\-lement\-izer-like items in Tsez viz. {\itshape {\l}i},
which permits long-distance agreement (and shows up in all Tsez
examples exhibiting long-distance agreement disussed above), and {\it
  \textcrlambda in}, which blocks long-distance agreement. As noted,
\cite{Polinsky&Potsdam:01} assume that {\itshape \textcrlambda in} is
indeed a regular C item, and assuming that the presence of a CP makes
head government impossible, it is correctly predicted that there is no
long-distance agreement across {\itshape \textcrlambda in}.
However,  for the same reason, it must be assumed that {\itshape {\l}i} is
not a C element. The problem here is that this is exactly what
it looks like, given that, like {\itshape \textcrlambda in}, it is the
outermost head in the word containing V. What is more, it does not yet suffice to
assume that  {\itshape {\l}i} not is not a C element -- {\itshape {\l}i} must  be
assumed not to be structurally represented {\itshape at all}. The reason is
that if {\itshape {\l}i} were the head of a phrase (outside of TopP), it would block
long-distance agreement in the same way as {\itshape \textcrlambda in}. 
It remains unclear whether there is any independent evidence for such
a radically different treatment (projecting complementizer
vs. structure-less morphological marker) of the two items. (In the
analysis to be developed in section 3 below, we will presuppose that
both  {\itshape \textcrlambda in} and {\itshape {\l}i} are regular C items.)

We take these first two problems to be potentially worrisome but
certainly not decisive. Arguably, things are different with the next
two issues raised by Polinsky \& Potsdam's analysis, concerning a
semantic problem based on the assumed covert DP movement, and an
incompatibility of the analysis with what look like clear cases of
long-distance agreement across a CP boundary.


\subsubsubsection{Topic Interpretation Within the Embedded Clause}

\cite{Polinsky&Potsdam:01} assume that the landing site of the
abstract movement is in the left periphery of the embedded
clause. Accordingly, the long-distance agreement-controlling DP is
interpreted as the topic of the embedded clause. A problem with this
analysis is that information structure phenomena -- and root phenomena
in general -- are usually confined to clauses that have some
illocutionary force; see \cite{Hooper&Thompson:73},
    \cite{Ebertetal08}, \cite{Krifkainprep}, and \cite{Maticetal14}. Most long-distance agreement-allowing
matrix verbs, however, are factives, which semantically take sentence
radicals (see \cite{Stenius67}), i.e.~propositions, as complements (see \cite{Krifka04})
 and also syntactically involve smaller
structures (see \cite{deCubaUrogdi10}). Thus, complements of
factives do not involve any illocutionary operator in their
syntax/semantics -- only the matrix clause does. Under a structured
proposition approach  (see \cite{Krifka92}) to information structural
phenomena, this leads to a semantic representation in which the topic
can only be understood as the topic of the whole sentence. That is,
the predicted structure for \ref{2-b}  (repeated in
\Next below) is as in (\NNext).

\ea\label{ex:mueller:13}
\gll    Eni-r [$\alpha$ u\v{z}\={a} \label{2d-b}magalu b-\={a}c'-ru-\l i~] b-iy-xo \\
     mother-{\scshape dat} {} boy-{\scshape erg} bread.{\scshape iii.abs} {\scshape iii}-eat-{\scshape pstprt-nmlz} {\scshape iii}-know-{\scshape prs} \\
\glt      `The mother knows that the boy ate the bread.'
\z

\ea\label{ex:mueller:14} {\scshape Assert}($<$$\_$T $\lambda$P.{\bfseries mother}($\lambda$z.{\bfseries boy}($\lambda$x.P($\lambda$y.eat(x, y))) ($\lambda$p.know(z, p))), {\bfseries bread}$>$)\z
The felicity conditions of the {\scshape Assert}-operator for topic-comment
structures make reference to the first part of the structured
proposition as a whole. Thus, it is unclear whether the embedded topic interpretation
advocated by \cite{Polinsky&Potsdam:01}  is actually available;
i.e.,  whether (\LLast)\ {\itshape is} actually understood as paraphrased in \ref{embeddedtopicint}, or not rather as in \ref{matrixtopicint}.
   
\ea\label{ex:mueller:15} Readings for topics in long-distance agreement
\ea \label{embeddedtopicint}The mother knows that, as for the bread, the boy ate it.
\ex \label{matrixtopicint}As for the bread, the mother knows that the boy ate it.
\z
\z
More specifically, complex sentences with factive matrix verbs
presuppose the truth of the proposition denoted by the respective
complement clause. In terms of information structure, factive
presuppositions belong to the ``background'' of an utterance and they
are ``taken for granted''. Thus, they are not ``at-issue'' or ``under
discussion''. This characterisation is inconsistent with
\citegen{Polinsky&Potsdam:01} assumption that a DP embedded under a factive verb
can act as topic of the embedded clause.

\subsubsubsection{Long-Distance Agreement Across a CP Boundary}

A severe empirical problem for the analysis developed in
\cite{Polinsky&Potsdam:01} (but also for analyses of the small
structure type) is that there is evidence from other Nakh-Daghestanian
languages that strongly suggests that long-distance agreement is in
principle possible across a CP boundary, and without movement to SpecC
(recall footnote \ref{whabs}). 

Thus, Khwarshi (see
\cite{Khalilova09}) and Hinuq (see \cite{Forker:11}), two
Nakh-Daghe\-stanian languages closely related to Tsez,
 also exhibit long-distance agreement. Similarly to Tsez, this
also goes along with a prominent information structural status of the
triggering DP. In contrast to Tsez, however, in these languages, the
triggering NP can have either topic or focus status.
For instance, in the Khwarshi examples in (\Next), the 
absolutive DP is interpreted as the topic of the embedded clause, and
long-distance agreement is possible.
However, long-distance agreement is also possible in answers to
information questions, as in (\NNext), suggesting that the
long-distance agreement-controlling DP$\_${abs} may also function as the focus of the embedded clause.

\ea\label{ex:mueller:16}
\ea
\gll I\v{s}et'u-l l-iq'-\v{s}e goli u\v{z}a bataxu y-acc-u \\
 mother.{\scshape obl-lat} IV-knows-{\scshape prs} {\scshape cop} boy.{\scshape erg} bread(V) V-eat-{\scshape pst.ptcp} \\
\glt `Mother knows that the boy ate bread.'
\ex  
\gll I\v{s}et'u-l \label{16-b}y-iq'-\v{s}e goli u\v{z}a bataxu y-acc-u \\
  mother.{\scshape obl-lat} V-knows-{\scshape prs} {\scshape cop} boy.{\scshape erg}  bread(V) V-eat-{\scshape pst.ptcp} \\
\glt `As for the bread, mother knows that  the boy ate it.'
\z
\z

\ea\label{ex:mueller:17}
\ea (Which cow does the boy know came?)
\ex
\gll  U\v{z}a-l l/b-iq'-\v{s}e \label{17-b}k\textsuperscript{\textipa{Q}}aba zihe b-ot'uq'q'-u \\
  boy.{\scshape obl-lat} IV/III-know-{\scshape prs} black cow(III) III-come-{\scshape pst.ptcp} \\
\glt `The boy knows that the black cow has come.'
\z
\z
Similar facts obtain in Hinuq.  

Importantly, long-distance agreement in Khwarshi and Hinuq is also less
restricted in another respect:
There are cases in which a {\em wh}-element occurs in
     an interrogative  complement clause, and long-distance agreement is nevertheless
     available. This is shown for Khwarshi in (\Next) (see \cite{Khalilova09}). 

\ea\label{ex:mueller:18}
\gll U\v{z}a-l l/b-iq'-\v{s}e [CP$\_${\rm (IV)} \l \label{cp1}u$\_${\rm [foc]} zihe b-iti-xx-u~] \\
boy.{\scshape obl-lat} {\scshape iv/iii}-know-{\scshape prs} {} who.{\scshape erg} cow({\scshape iii}) {\scshape iii}-divide-{\scshape caus-pst.ptcp} \\
\glt `The boy knows who has stolen the cow.'
\z
Here a wh-phrase bearing ergative case shows up in the embedded
interrogative clause. Given standard assumptions about the
semantics of questions (see, e.g., \cite{Stechow:96:aga}),
the interrogative interpretation of a clause is inherently, and
invariably, tied to the presence of a C element. Therefore, (\Last)
proves that long-distance agreement across a CP boundary is possible
in Khwarshi independently of whether the ergative wh-phrase can be
assumed to be located in SpecC in the syntax, and of whether or not
there is an overt C item present. 

Next, (\Next) and (\NNext) illustrate the possibility of long-distance
agreement across a CP boundary in Hinuq (see \cite{Forker:11}). In
(\Next), the embedded interrogative clause contains a subject
wh-phrase that does not block long-distance agreement with the
(non-wh) absolutive DP. 

 \ea\label{ex:mueller:19} 
\gll\relax [CP \textbeltl u rek'$^{w}$e go\textbeltl \label{cp2}i\v{s}~] di\v{z} \o -eq'i-yo gom \\
  {} who man.{\scshape i} be.{\scshape cvb} I.{\scshape dat} {\scshape i}-know-{\scshape prs}  be.{\scshape neg} \\ 
\glt `I do not know who he was.'
\z
In contrast, (\Next) shows that long-distance agreement is also possible with
an absolutive DP that is a wh-element itself; but there is no reason
to assume that DP$\_${abs} is not in its base position here. 

\ea\label{ex:mueller:20} \label{cp3}
  \gll Debez r/\o -eq'i-ye [CP(V) k'a\v{c}a\textgamma -za-y \textbeltl u \o -uher-i\v{s}-\textbeltl i~]~? \\
  you.{\scshape sg.dat} {\scshape v/i}-know-{\scshape q} {} bandit-{\scshape obl.pl-erg}   who {\scshape i}-kill-{\scshape res-abst} \\ 
 \glt `Do you know whom the bandits killed?'
\z
Note that (\LLast) and (\Last) contain clause-final elements that
would seem to correspond to {\itshape {\l}i} rather than  {\it
  \textcrlambda in} in Tsez. However, notwithstanding the problems
raised by \citegen{Polinsky&Potsdam:01} analysis of {\itshape {\l}i}
mentioned above, and notwithstanding the arguments that Forker
presents for a uniform CP analysis of these contexts in Hinuq after
all, it is clear that the status of {\itshape {\l}i} has no bearing on the
question of whether there is a CP present in \ref{cp2} and \ref{cp3}
(and \ref{cp1}, for that matter): There must be a CP boundary here
because of the combination of interrogative semantics and a full
clausal structure, including assignment of ergative (which by itself
is not yet decisive, given that there are good arguments for assuming
that the ergative is assigned within vP in Nakh-Daghestanian
languages; see \cite{Gagliardietal:14}, \cite{Polinsky:16:arc}). 

Thus, there is strong evidence from Khwarshi and Hinuq for the general
availability of long-distance agreement across what must qualify as a
CP.

What is more, unlike Tsez, the non-Tsezic Nakh-Dahgestanian language
Tsakhur  also permits what can be called
``super-long-distance agreement'', i.e., long-distance agreement 
across two clause boundaries; see \cite{Kibrik:99}.

\ea\label{ex:mueller:21} \label{tsa1}
\gll I\v{c}\={\textsci}-s w=u\={k}\={\textsci}k\textbari n-na [CP ji\v{c}o-j-s [CP gaba  h\={a}\textglotstop-as~] XaIr-qi=w=x-es~] \\
     girl-\textsc{dat} {3}=want.\textsc{pf-{aa}} {} sister-\textsc{obl-dat} {} {carpet.3} 3.do-\textsc{pot} learn-{3}=become-\textsc{pot} \\
\glt `The girl wants her sister to learn to make a carpet.'
\z
Similar phenomena have also been reported by \cite{Forker:11} for
Hinuq; see (\Next).

\ea \label{ex:mueller:22}\label{tsa2}
\gll Iyo-z b-eq$'$i-yo [CP Pat$'$imat-ez [CP tort  b-ac$'$-a~] b-eti-\v{s}-\l i~]]\\
mother-{\scshape dat} III-know-{\scshape prs} {} Patimat-{\scshape dat} {} cake(III) III-eat-{\scshape inf} III-want-{\scshape res}-{\scshape abst}\\
\glt `The mother knows that Patimat wanted to eat the cake.''
\z
If there are two CP boundaries present, there is no way for the feeding approach
to account for the option of long-distance agreement since the covert
movement postulated in this approach always has to be
clause-bound.

\subsection{Interim Conclusion}

We take it that the conclusion that can be drawn on this basis is that
all four existing approaches face significant problems with
long-distance agreement from the point of view of a grammar that
incorporates a strict locality principle like the PIC. Thus, there is
every reason to pursue a  new approach; and given the general
availability of long-distance agreement across a CP, this approach
must be such that it preserves strict locality even if there can be no
denying the fact that the matrix verb and the embedded agreement
controller DP can be far away from one another in structural terms in
syntactic surface representations. 

As a basic premise, we will assume that the only way to locally model
non-local dependencies is via {\itshape movement} (see
\cite{Hornstein:01,Hornstein:09} for this general point, and
\cite{Mueller:14:buf} for some specific proposals in a priori
recalcitrant domains). Given the PIC, a matrix V and an embedded
(agreement-controlling) DP have to enter a local relation at some
point of the derivation. As we have seen, there is evidence against
the assumption that DP moves to the matrix V domain (or to a position
of the embedded domain that is accessible from it); this excludes feeding
analyses. The only remaining possibility then is that it is actually
V that moves to the matrix domain: If the mountain won't come to the
prophet, the prophet will go to the mountain.

For concreteness, we would like to propose that locality in
long-distance agreement is not established {\itshape late} in the
derivation (as \citegen{Polinsky&Potsdam:01} approach, where
movement feeds long-distance agreement), but, in fact, {\it
  early}. This approach thus involves counter-bleeding (rather than
feeding): Agreement with the embedded internal argument DP takes place
at a stage in the derivation when DP and the two verbs involved are
all clause-mates. It is only due to subsequent reprojection movement
of what will eventually become the matrix verb 
 that on the surface it looks as if agreement takes place
long-distance; reprojection movement of V thus comes too late to bleed
(i.e., it counter-bleeds) Agree.


\section{A New Analysis}

\subsection{Head Movement as Reprojection}

Let us begin by sketching the outlines of a general approach to head
movement in terms of reprojection, a concept that has been widely
pursued for various empirical domains over the last decades (see
\cite{Pesetsky:85}, \cite{Stechow&Sternefeld:88},
\cite{Sternefeld:89}, \cite{Holmberg:91}, \cite{Ackemaetal:93},
\cite{Kiss:95}, \cite{Koeneman:00}, \cite{Haider:00:bra},
\cite{Bhatt:02}, \cite{Hornstein&Uriagereka:02},
\cite{Fanselow:03,Fanselow:09:boo}, \cite{Bury:03}, \cite{Suranyi:05},
\cite{Donati:06}, \cite{Bayer&Brandner:08},
\cite{Georgi&Mueller:10:rep}, \cite{Mueller:11:loc}, and
\citetalias{SMueller:15}, among others).\footnote{\cite{Pesetsky:85}
  suggests that reprojection after head movement at LF serves to
  circumvent bracketing paradoxes. As far as we can tell, this
  qualifies as the first instance of a reprojection approach to head
  movement in the literature.} The basic idea behind head movement as
reprojection is that an X$^0$ head is moved out of a projection that dominates it 
and takes this projection as its own complement by merging with it,
  projecting anew in the derived position. This solves the notorious
  c-command and Extension Condition (cf. \cite{Chomsky:95}) problems with head movement as
  adjunction to an X$^0$ category: In a head-movement-as-adjunction structure like (\Next),
the moved head Y fails to extend the tree (since XP must, be
definition, have been in place before movement of Y), and Y does not
c-command its trace (because the next branching node containing Y is
the higher X segment, which does not dominate Y's trace).

\ea\label{ex:mueller:23}\relax [XP [X Y$\_$1 X~] [WP ... t$\_$1 ...~]]\z
These problems disappear under a reprojection approach:  Head movement has now extended the
  tree, and the moved item is able to c-commands its
  trace. Furthermore, this approach does not 
  necessitate (i) a relocation of head movement to PF (see
  \cite{Chomsky:00}), (ii) a
  reinterpretation as XP movement (see \cite{Koopman&Szabolcsi:00}, \cite{Mahajan:01},
  and \cite{Nilsen:03:diss}, among many others), or (iii) the
  postulation of a complex
  operation integrating both regular syntactic movement and
  syntactically irregular morphological merger (see
  \cite{Matushansky:06}). 

There are basically three different reprojection scenarios. A first
possibility is that a head moves out of its own projection, merges
with the XP of which it was the head prior to the movement, and
projects anew. Such local reflexive reprojection is shown in (\Next). 

\ea\label{ex:mueller:24} {\itshape Local reflexive reprojection}:\\
%\begin{tabular}{ccccccc}
% & \node{0}{XP}&&&&\\[4mm]
%\node{1}{X$\_$1} && \node{2}{XP} \\[4mm]
%& \node{3}{WP} && \node{4}{X$'$}\\[4mm]
%&&\node{5}{t$\_$1} && \node{6}{ZP}\\[4mm]
%\nodeconnect{0}{1}\nodeconnect{0}{2}\nodeconnect{2}{3}\nodeconnect{2}{4}
%\nodeconnect{4}{5}\nodeconnect{4}{6}
%\anodecurve[b]{5}[b]{1}{2cm}
%\end{tabular}
\z
A second possibility is that a reprojection movement is still highly
local (in the sense that the moved head attaches to the minimal phrase
that dominated it before the movement step was carried out), but not
reflexive. In this scenario, the moved head excorporates from a complex
head structure that was formed by an earlier (possibly pre-syntactic) 
operation combining two primitive X$^0$ categories (in accordance with
c-command and Extension Condition requirements), or that is stored as
such in the lexicon; after the movement, the moved head projects its own
XP in the derived position.\footnote{Thus, strictly speaking, this is
  not actually an instance of {\itshape re}-projection: X in \ref{aux12}
  projects for the first time in the derived position.} Local non-reflexive
reprojection is illustrated in (\Next). 

\ea\label{ex:mueller:25} {\itshape Local \label{aux12}non-reflexive reprojection}:\\
%\begin{tabular}{ccccccc}
% & \node{0}{XP}&&&&\\[4mm]
%\node{1}{X$\_$1} && \node{2}{YP} \\[4mm]
%& \node{3}{WP} && \node{4}{Y$'$}\\[4mm]
%&&\node{5}{Y-t$\_$1} && \node{6}{ZP}\\[4mm]
%\nodeconnect{0}{1}\nodeconnect{0}{2}\nodeconnect{2}{3}\nodeconnect{2}{4}
%\nodeconnect{4}{5}\nodeconnect{4}{6}
%\anodecurve[br]{5}[b]{1}{2cm}
%\end{tabular}
\z
Finally, reprojection can be non-local (by definition, it is then also
non-reflexive). In (\Next), the moved head skips over two maximal
projections and reprojects in the derived position. 

\ea\label{ex:mueller:26}{\itshape Non-local \label{25}non-reflexive reprojection}:\\
%\begin{tabular}{ccccccc}
% & \node{0}{XP}&&&&\\[4mm]
%\node{1}{X$\_$1} && \node{2}{YP} \\[4mm]
%& \node{3}{WP} && \node{4}{Y$'$}\\[4mm]
%&&\node{5}{Y} && \node{6}{XP}\\[4mm]
%&&&\node{7}{ZP} && \node{8}{X$'$} \\[4mm]
%&&&&\node{9}{t$\_$1} && \node{10}{UP}\\
%\nodeconnect{0}{1}\nodeconnect{0}{2}\nodeconnect{2}{3}\nodeconnect{2}{4}
%\nodeconnect{4}{5}\nodeconnect{4}{6}
%\nodeconnect{6}{7}\nodeconnect{6}{8}
%\nodeconnect{8}{9}\nodeconnect{8}{10}
%\anodecurve[b]{9}[b]{1}{2cm}
%\end{tabular}
\z
Assuming these three scenarious to be available in the world's
languages, it can be concluded that head movement can involve
excorporation (see \cite{Roberts:91,Roberts:97:res}), and that head
movement does not obey the Head Movement Constraint (see
\cite{Roberts:09:hea,Roberts:10} vs. \cite{Travis:84} for arguments to
this effect). Given that the data that originally motivated
stipulation of the excorporation and Head Movement Constraint
restrictions can be derived otherwise, this would seem to permit a
simpler, more attractive theoretical approach, and to correspond to
the null hypothesis. Furthermore, one should expect that head movement
as reprojection obeys the same constraints that hold of all movement
operations; this includes the PIC (see \ref{pic1}). Thus, for the
operation to be legitimate, it can be concluded that YP is not a phase
in (\Last); and that, more generally, head movement as reprojection can
cross phases by carrying out intermediate movement steps to phase
edges, in accordance with the PIC.

As for the concrete mechanics of reprojection movement, we will make
the following assumptions. First,  all syntactic operations are
feature-driven: On the one hand, there are designated   structure-building features (edge features, subcategorization
features) that trigger (external or internal) Merge; we will refer to
these as [{\small $\bullet$}F{\small $\bullet$}] features. On the
other hand, there are   probe features trigger Agree. To simplify
exposition and simultaneously avoid commitment to one of the existing options in
various domains (e.g., valuation vs. checking, interpretability
vs. uninterpretability), we will refer to probes as 
 [$*$F$*$] features throughout.
All these features triggering syntactic Merge and Agree operations are
ordered on lexical items; and they are discharged (i.e., rendered
syntactically inactive) one after the other
after having induced the respective operations that they encode. 
Finally (although this assumption will not actually be crucial), we
postulate that all phrases are phases. As a consequence, movement must take place via all
  intermediate phrase edges  that intervene between a base position and
  the ultimate landing site of some moved item (except for the minimal
  specifier domain if the item is already part of the phase edge, as
  is the case with reprojection movement of heads). Given this
  assumption, YP in \ref{25} must be a phase, and X$\_$1 must therefore
  carry out an intermediate step to SpecY on its way to its ultimate
  position.\footnote{One might think that allowing heads to move to
    specifier positions might give rise to various over-generation
    problems. However, it is worth bearing in mind that specifier
    positions can only ever be used as {\itshape intermediate} escape
    hatches (required by the PIC) by head movement under present
    assumptions; thus, the situation is completely analogous to, say,
    wh-movement via an intermediate Specv position in English -- the
    wh-phrase can use this position as an intermediate escape hatch,
    but can never ultimately show up in it (for essentially the same
    reason, viz., that the trigger can only saturated in the final
    landing site).}

Suppose further that Featural Cyclicity holds, as in
(\Next).\footnote{This constraint can plausibly be derived as a theorem
  under various conceptions of cyclic spell-out of complements of
  phase heads.}

\ea\label{ex:mueller:27} {\itshape Featural \label{fc}Cyclicity}:\\
A non-root XP cannot contain a feature $\Gamma$ in the non-edge domain of X that is supposed to trigger an operation ([{\small $\bullet$}F{\small $\bullet$}] or [$*$F$*$]).\z
In the normal course of events, the head X of some XP has discharged
all the Merge-inducing features ([{\small $\bullet$}F{\small $\bullet$}])
 and Agree-inducing features ([$*$F$*$]) it contains before XP is merged with some
 other category. However, suppose that the head X has not been able to
 discharge a [{\small $\bullet$}F{\small $\bullet$}] or [$*$F$*$]
 (plus, possibly, other features that are lower on the list of
 operation-inducing features of the head, and that can only be accessed
 if the topmost feature has been discharged). In such a situation, one
 of two Last Resort operations may take place: Either the [{\small
     $\bullet$}F{\small $\bullet$}] or [$*$F$*$] feature is deleted (see 
\cite{Bejar&Rezac:09},
  \cite{Preminger:14}, and  \cite{Georgi:14} for proposals along these
  lines); or the item containing the incriminating feature is moved to the edge domain of
  the current phrase, so as not to violate Featural Cyclicity
  in (\Last). The two Last Resort options for [{\small
      $\bullet$}F{\small $\bullet$}] and [$*$F$*$] features are stated
  in (\Next). 

\ea\label{ex:mueller:28} {\itshape Last Resort}:\\
If a feature $\Gamma$ on X that triggers an operation cannot be discharged
in XP, there are two basic options:
\ea $\Gamma$ is deleted.
\ex $\Gamma$ is moved to the edge of XP, pied-piping the minimal
category containing it. 
\z
\z
Thus, a head X with a non-discharged $\Gamma$ ([{\small
    $\bullet$}F{\small $\bullet$}] or [$*$F$*$]) feature undergoes
intermediate movement to phrase edges for as long as it takes to reach
a position in which $\Gamma$ can eventually be discharged.  Following
\cite{Fanselow:03,Fanselow:09:boo}, \cite{Suranyi:05},
\cite{Matushansky:06}, and \cite{Georgi&Mueller:10:rep}, these kinds
of features can then be viewed as triggers for reprojection
movement.\footnote{\cite{Fanselow:03,Fanselow:09:boo} and
  \cite{Georgi&Mueller:10:rep} refer to these kinds of features as
       {\itshape M\"unchhausen} features, based on the literary character
       Baron M\"unchhausen who escapes from a swamp (where he is
       trapped on the back of his horse) by pulling himself up by his
       hair.} Note that it can in principle be both probe features on
some head X that trigger (intermediate or final) reprojection movement
(e.g., if there is no matching goal for a probe in the structure, or
if the goal is not c-commanded by the probe feature on X), and
structure-building features (e.g., if there is no accessible matching
category, or if two heads simulteneously need to discharge their
[{\small $\bullet$}F{\small $\bullet$}] feature but only one can do
this at any given stage of the derivation). However, in the
reprojection approach to long-distance agreement to be developed in
the next section, it is the need to discharge a structure-building
feature that triggers the movement of (what thereby becomes) the
matrix verb.


\subsection{Long-Distance Agreement by Reprojection}

\subsubsection{Complex Predicates}

Long-distance agreement typically encompasses verbs that in many
languages are restructuring verbs.  In fact, for another
Nakh-Daghestanian language, Godoberi, \cite{Haspelmath99} shows with a
series of tests that apparent long-distance agreement in the language
actually involves only a monoclausal structure with a complex
predicate.  However, \citet{Forker:11} and \citet{Khalilova09} show
with similar tests that this is not the case for Hinuq or Khwarshi,
both of which involve truly biclausal structures, with an embedded CP.

In view of this state of affairs, we would like to suggest that
despite this biclausal character, long-distance agreement in Hinuq and
Khwarshi (and Tsez, and perhaps more generally) does indeed involve
some form of restructuring, albeit in the form of a special type of
complex predicate formation. In standard lexical approaches to complex
predicate formation (see, e.g., \cite{Haider:93,Haider:10},
\cite{Kiss:95}, \cite{Stiebels:96}, and \citetalias{SMueller:02} on
German, or \cite{Butt:95} on Hindi/Urdu), all lexical
subcategorization information of the verbs that participate in the
operation is unified by functional composition.  This results in one
featural array for the complex predicate and monoclausality
throughout. Against the background of the present approach, this would
imply a unique list of structure-building and probe features
associated with the complex predicate, with the features discharged
one after the other. In contrast, we adopt a version of pre-syntactic
complex predicate formation where two predicates (two verbs, in the
case at hand) are combined into a complex category in a way that, crucially,
leaves the verbs' individual lexical information intact -- i.e., there
are still two separate lists of features triggering syntactic
operations.\footnote{For present purposes, it is immaterial whether
  this pre-syntactic component is conceived of as the lexicon, or as a
  pre-syntactic morphology domain; for concreteness, we will generally assume
  the former here.}

\subsubsection{Derivations}

Let us now look at how long-distance agreement in Nakh-Dahestanian
languages (and possibly elsewhere) can be derived on the basis of an
approach in terms of reprojection and pre-syntactic complex predicate
formation. Throughout, we will assume a CP status of the embedded
clause, with both  {\itshape \textcrlambda in}-type and {\itshape {\l}i}-type
markers qualifying as C heads. The definition of Agree that we will
adopt is similar but not identical to the one in \ref{Agree} above; it
is given in (\Next).\footnote{Since the PIC holds for all syntactic
  operations, the fact that Agree is also subject to this constraint
  does not have to be mentioned explicitly. As noted above, we will
  not address the question here of how exactly other cases of Agree that
  would at first sight seem to violate the PIC can be accounted for;
  but note that this issue is even more prominent (though not
  categorially different) in an approach where all phrases are
  phases. 

The requirement in \ref{agua2} permits both upward and downward Agree; see
\cite{Zeijlstra:12} and \cite{Bjorkman&Zeijlstra:14} vs. \cite{Preminger:13:tha}.
(The local Agree operation initiated by (what will become) the matrix verb in
long-distance agreement will involve upward Agree.) \ref{agua3}
ensures minimality, with closeness definable in terms of minimal path
length. There is no defective intervention here: Discharged
features on intervening  heads and checked features on intervening phrases can be ignored. Finally,
\ref{agua4} encodes the Activity Condition: An active feature is one
that has not participated in Agree.}

\ea\label{ex:mueller:29} {\itshape Agree}:\\
$\alpha$ can Agree with $\beta$ if (a)--(d) hold.
\ea $\alpha$ carries a probe feature [$*$F$*$], and $\beta$ carries a
maching goal feature [F].
\ex $\alpha$ \label{agua2}c-commands $\beta$, or $\beta$ c-commands $\alpha$.
\ex There is \label{agua3}no $\delta$ that is closer to $\beta$ than $\alpha$ and
also carries [$*$F$*$], and there is no $\gamma$ that is closer to
$\alpha$ than $\beta$ and carries an active [F].
\ex $\beta$ bears an \label{agua4}active feature. 
\z
\z
The syntactic derivation of a sentence such as \ref{cp3} in Hinuq,
where the matrix verb undergoes long-distance agreement with the
embedded absolutive wh-phrase, starts with the complex predicate in
(\Next); \ref{cp3} is repeated here as (\NNext) (with the default
agreement option ignored).\footnote{\label{vari2}One may ask why it is
  that V$\_$2 (which will eventually become the embedded verb) projects
  in this structure, rather than V$\_$1 (which will end up as the matrix
  verb).  As a matter of fact, there does not seem to be a good reason
  why the alternative representation where V$\_$1 projects -- {\it
    [$\_${V$\_$1}[$\_${V$\_$1}know]-[$\_${V$\_$2}kill]]} -- should be excluded as such:
  With two bare X$^0$ heads forming a complex structure, labelling can
  be expected to be free. However, as will become clear when we look
  at the derivation for \ref{cp3d}, choosing an initial representation
  {\itshape [$\_${V$\_$1}[$\_${V$\_$1}know]-[$\_${V$\_$2}kill]]} of the complex predicate
  (rather than {\itshape [$\_${V$\_$2}[$\_${V$\_$1}know]-[$\_${V$\_$2}kill]]}) can never lead
  to a well-formed derivation: V$\_$2 ultimately needs to combine with a
  DP, but after (extended) projection of V$\_$1 has been completed, only
  a CP is available for V$\_$2, and this makes it impossible to discharge
  the [{\small $\bullet$}D{\small $\bullet$}] feature of V$\_$2. In the
  same way, V$\_$1 needs to combine with a CP (due to [{\small
      $\bullet$}C{\small $\bullet$}] on its feature list), but such a
  CP is not available at the beginning of the derivation.}

\ea\relax\label{ex:mueller:30} [V$\_$2 [V$\_$1 know] [V$\_$2 kill]]\z

 \ea\label{ex:mueller:31}  \label{cp3d}
 \gll  Debez \o -eq'i-ye [CP(V) k'a\v{c}a\textgamma -za-y \textbeltl u \o -uher-i\v{s}-\textbeltl i~]~? \\
  you.{\scshape sg.dat} {\scshape i}-know-{\scshape q} {} bandit-{\scshape obl.pl-erg}   who {\scshape i}-kill-{\scshape res-abst} \\ 
 \glt `Do you know whom the bandits killed?'
\z
In the first step, {\itshape [$\_${V$\_$2}[$\_${V$\_$1}know]-[$\_${V$\_$2}kill]]} is merged
with  the internal argument DP, triggered by  [{\small
    $\bullet$}D{\small $\bullet$}] on V$\_$2. The resulting
representation is shown in \ref{Baum}.

\ea\label{ex:mueller:32} {\itshape Long-distance agreement \label{Baum}by reprojection, first stage}:
%\begin{tabular}{ccccccccccccccc}
%&&&&&&& && \node{16}{VP}\\[4mm]
%&&&&&&&&\node{17}{V$\_$2} && \node{18}{DP$\_${abs}}\\[4mm]
%&&&&&&&\node{19}{V$\_${1,[\bullet \rm C\bullet]}} && \node{19a}{V$\_$2} &\node{20}{[$\phi$]} \\
%&&&&&&&\node{21}{[*$\phi$*]} && \node{22}{[*$\phi$*]} & \node{23}{[inf-st]}\\
%&&&&&&&\node{24}{[*inf-st*]} && \\
%\end{tabular}
%\nodeconnect{16}{17}\nodeconnect{16}{18}
%\nodeconnect{17}{19}\nodeconnect{17}{19a}\nodeconnect{18}{20}
\z

Each of the verbs involved has its own $\phi$-probe
(see \cite{Bejar&Rezac:09}), which is checked through Agree with the
$\phi$-feature on the internal argument DP. Since V$\_$2 is the head of
the complex predicate, its $\phi$-probe intervenes between V$\_$1's
$\phi$-probe and the internal DP.  Thus, V$\_$2's $\phi$-probe has to be
discharged first; afterwards, V$\_$1 can discharge its  $\phi$-probe via
Agree with DP. This
derives the generalization that long-distance agreement (i.e., under
present assumptions, extremely local agreement of V$\_$1 and the absolutive DP)
is possible only if embedded agreement (i.e., agreement of V$\_$2 and the
absolutive DP) has taken place.\footnote{Given that only absolutive
  DPs can act as agreement controllers in the languages currently
  under consideration, the question arises whether this information is
  already locally available in the structure in \ref{Baum}. There are
  two possibilities, both of which strike us as viable. First, the
  absolutive (vs. lexically case-marked) nature of an internal DP
  argument might indeed already be visible at this stage (e.g.,
  because V$\_$2 does not have a lexical case feature). Second, if
  absolutive is not identifiable yet at this stage, agreement could
  simply take place in the hope that it will later emerge (e.g., be
  assigned by a functional head like t) -- if it does not, the
  derivation will eventually crash.}

In addition to the dependence of long-distance agreement on local
agreement, a second generalization about long-distance can be derived
at this point: There must be an obligatory information-structural
reflex on the DP participating in long-distance agreement (with an
interpretation as topic in Tsez, as topic or focus in Hinuq and
Khwarshi, etc.); this is simply signalled by [inf-st] in (\Last) (i.e.,
[inf-st] stands for [topic], [topic, focus], or other
information-structural features). Here is why: 
Given  \citegen{Chomsky:01} {\itshape Activity Condition}, after $\phi$-Agree
  with V$\_$2, DP$\_${abs} in \ref{Baum} can only undergo $\phi$-Agree
  with V$\_$1 if it still has {\itshape a different}, {\itshape active} feature
  that V$\_$1 is looking for; and [inf-st] fulfills this role.
 This explains the presence of [*inf-st*] on V$\_$1 that needs to undergo
 Agree with [inf-st] on  DP$\_${int}. As a consequence of this second
 Agree operation involving V$\_$1 and DP$\_${int}, V$\_$1 is equipped with the
 information that DP$\_${int} is a topic. 

 In the further course of the derivation, V$\_$2 first discharges all its
 structure-building features (if it has any such featuress
 left). Subsequently, v merges with VP; after that it merges 
 with an external argument DP; and then it assigns ergative case to it. At
 this point, V$\_$1 has not yet had a chance to discharge its [{\small
    $\bullet$}C{\small $\bullet$}] feature.\footnote{This
   feature is either lower on the list of operation-triggering
   features of V$\_$1 than the probe features for agreement with
   DP$\_${int}, or there are actually two separate stacks involved here
   (as indicated in \ref{Baum}):
   one for structure-building features, and one for probe
   features. This second option might be preferable on conceptual and empirical
   grounds; see \cite{Mueller:04:arg,Mueller:09:eao} for discussion.}
  Therefore, before
 the vP is completed, V$\_$1 needs to move to v's specifier position, so
 as to comply with Featural Cyclicity (cf. \ref{fc}). The resulting
 representation is shown in (\Next).

\ea\label{ex:mueller:33} {\itshape Long-distance agreement \label{Baum2}by reprojection, second
stage}:\\
%\begin{tabular}{ccccccccccccccc}
%&&&&&& \node{10}{vP}\\[4mm]
%&&&&&\node{11}{V$\_${1,[\bullet \rm C\bullet]}} && \node{12}{v$'$}\\[4mm]
%&&&&&&\node{13}{DP$\_${erg}} && \node{14}{v$'$}\\[4mm]
%&&&&&&&\node{15}{v} && \node{16}{VP}\\[4mm]
%&&&&&&&&\node{17}{V$\_$2} && \node{18}{DP$\_${abs}}\\[4mm]
%&&&&&&&\node{19}{t$\_${V$\_$1}} && \node{19a}{V$\_$2} &\node{20}{[$\phi$]} \\
%&&&&&&&\node{21}{[*$\phi$*]} && \node{22}{[*$\phi$*]} & \node{23}{[inf-st]}\\
%&&&&&&&\node{24}{[*inf-st*]} && \\
%\end{tabular}
%\nodeconnect{10}{11}\nodeconnect{10}{12}
%\nodeconnect{12}{13}\nodeconnect{12}{14}\nodeconnect{14}{15}\nodeconnect{14}{16}\nodeconnect{16}{17}\nodeconnect{16}{18}
%\nodeconnect{17}{19}\nodeconnect{17}{19a}\nodeconnect{18}{20}
%\anodecurve[l]{19}[b]{11}{2cm}
\z
In further steps, the TP and CP structures of (what will become) the
embedded clause are generated by Merge and Agree operations, while V$\_$1 moves up
   the developing syntactic structure, via intervening phase
   edges. Finally, when the CP is completed, and V$\_$1 has moved to C's
   edge domain because it still has not been able to discharge its
   structure-building feature [{\small
    $\bullet$}C{\small $\bullet$}], V$\_$1 is in a position from which it
  can undergo reprojection movement, take the CP generated so far as
  its complement (thereby discharging [{\small
    $\bullet$}C{\small $\bullet$}]), and create a matrix
  VP.\footnote{In addition to subcategorization, V$\_$1 carries out an
    Agree operation with C that reflects the embedding of an
    interrogative ([+wh]) clause.} This is
  shown in (\Next).\footnote{The mechanics here are similar to 
   \citegen{Martinovic:15} analysis of the left periphery of Wolof
   in terms of head splitting and reprojection.}   

\ea\label{ex:mueller:34} {\itshape Long-distance agreement \label{Baum3}by reprojection, third stage}:
%\begin{tabular}{ccccccccccccccc}
%& \node{0}{VP}\\[4mm]
%\node{1}{V$\_$1} && \node{2}{CP}\\[4mm]
%& \node{3}{~~} && \node{4}{C$'$}\\[4mm]
%&& \node{5}{C} && \node{6}{TP}\\[4mm]
%&&&\node{7}{~~} && \node{8}{T$'$}\\[4mm]
%&&&&\node{9}{T} && \node{10}{vP}\\[4mm]
%&&&&&\node{11}{~~} && \node{12}{v$'$}\\[4mm]
%&&&&&&\node{13}{DP$\_${erg}} && \node{14}{v$'$}\\[4mm]
%&&&&&&&\node{15}{v} && \node{16}{VP}\\[4mm]
%&&&&&&&&\node{17}{V$\_$2} && \node{18}{DP$\_${abs}}\\[4mm]
%&&&&&&&\node{19}{t$\_${V$\_$1}} && \node{19a}{V$\_$2} &\node{20}{[$\phi$]} \\
%&&&&&&&\node{21}{[*$\phi$*]} && \node{22}{[*$\phi$*]} & \node{23}{[inf-st]}\\
%&&&&&&&\node{24}{[*inf-st*]} && \\
%\end{tabular}
%\nodeconnect{0}{1}\nodeconnect{0}{2}\nodeconnect{2}{3}\nodeconnect{2}{4}\nodeconnect{4}{5}\nodeconnect{4}{6}
%\nodeconnect{6}{7}\nodeconnect{6}{8}\nodeconnect{8}{9}\nodeconnect{8}{10}\nodeconnect{10}{11}\nodeconnect{10}{12}
%\nodeconnect{12}{13}\nodeconnect{12}{14}\nodeconnect{14}{15}\nodeconnect{14}{16}\nodeconnect{16}{17}\nodeconnect{16}{18}
%\nodeconnect{17}{19}\nodeconnect{17}{19a}\nodeconnect{18}{20}
%\anodecurve[l]{19}[b]{11}{2cm}
%\anodecurve[l]{11}[b]{7}{1cm}
%\anodecurve[l]{7}[b]{3}{1cm}
%\anodecurve[l]{3}[b]{1}{.5cm}
\z
The resulting representation is opaque in \citegen{Kiparsky:73:abs}
sense as it involves a counter-bleeding interaction of operations
(also cf.  \cite{Chomsky:51}, \citet[25--26]{Chomsky:75:the}):
Reprojection movement of V$\_$1 would bleed Agree with DP$\_${abs}
(which requires strict locality, due to the PIC) but fails to do so
because it applies too late: When V$\_$1 has left the local domain in
which agreement  with DP$\_${abs} can legitimately be carried out, this
agreement has already taken place. 

From this point onwards, everything happens exactly as one would
expect it to (with matrix vP, TP, and CP generated by Merge and Agree
operations), and there is basically no difference anymore to
derivations in which there is no complex predicate formation to begin
with.  Of course, given that pre-syntactic (lexical) complex predicate
formation is an optional process, this second kind of derivation can
be assumed to underlie minimally different sentences in which
long-distance agreement does not occur. Thus, the two strategies
differ substantially as far as earlier stages are concerned, but they
end up with exactly the same structures once the matrix domain has
been reached. There is one qualification, though. 
 As a consequence of reprojection movement of V$\_$1, [inf-st] of
 DP$\_${int} is
  transported into the matrix clause.\footnote{Note that this implies
    that {\itshape discharged} features, while syntactically inert, are not
    actually deleted. This assumption
    must independently be made for
    discharged probe features more generally that give rise to
    morphological realization; see \cite{Adger:03}.} 
The information that the embedded DP$\_${\rm [inf-st]}
  is interpreted as a topic is therefore shifted to the matrix
  sentence, and consequently, a DP that is affected by long-distance
  agreement is interpreted as the topic of the entire complex sentence.
 The analysis is thus consistent with usual assumptions concerning the impossibility
  of information-struc\-tural elements in clauses without
  illocutionary force (like non-assertive, presuppositional
  declarative clauses). Whereas information-structural features of an
  embedded DP can thus be interpreted in the matrix clause, there is
  no way how an embedded DP could take relative scope in the matrix clause as
  well (cf. the sentences in \ref{scope32}): Relative scope is
  determined by the position of a item, not by features, and there is no stage of
  the derivation where the embedded DP$\_${abs} would show up in the
  matrix clause. 

\subsubsection{Further Consequences}

The example of long-distance agreement that we have considered here on
the basis of the sample derivation in \ref{Baum}, \ref{Baum2} and
\ref{Baum3}, involves a DP$\_${abs} controller that is also a
wh-phrase. However, it should be clear that the approach generalizes
to all the other cases of long-distance agreement mentioned
above. For instance, an account in terms of reprojection of the part
of a complex predicate straightforwardly derives long-distance
agreement as in \ref{2} in Tsez and in \ref{3} in Hinuq (where it can
now be assumed that $\alpha$ stands for a full CP). Similarly,
examples like \ref{16-b} in Khwarshi (with a DP$\_${abs} controller
acting as a topic, i.e., [inf-st] representing [topic]) and \ref{17-b}
in Hinuq (with a DP$\_${abs} controller acting as a focus, i.e., [inf-st]
representing [focus]) are directly accounted for under the present
analysis. \ref{cp1} (from Khwarshi) and \ref{cp2} (from Hinuq) have
subject wh-phrases (one marked by ergative, one not) that do not block
long-distance agreement with the absolutive DP. Again, this is
expected under present assumptions: Independently of whether the
wh-phrase here occupies SpecC in overt syntax or not, reprojection
movement of the verb to the matrix domain is possible (given the
general option of multiple specifiers, particularly for intermediate
movement steps). 

Next, instances of of super-long-distance agreement where the
agreeing verb and the agreement controller DP are separated by two
intervening CP boundaries, like \ref{tsa1} in Tsakhur or \ref{tsa2} in
Hinuq, can also be addressed under the reprojection approach: Here a complex
predicate is formed pre-syntactically where V$\_$1 (which will become the
highest verb) and V$\_$2 (which will become the intermediate verb) are
first combined, with V$\_$2 projecting (in a successful derivation;
cf. footnote \ref{vari2}), and then the complex V$\_$2 category is
combined with V$\_$3 (which will become the most deeply embedded verb),
with V$\_$3 projecting, as shown in (\Next).

\ea\relax\label{ex:mueller:35} [V$\_$3 [V$\_$2 V$\_$1 V$\_$2~] V$\_$3~]\z
Here V$\_$3, V$\_$2, and V$\_$1 first carry out Agree operations with V$\_$3's
internal argument (DP$\_${abs}), and then a CP is generated on top of
VP$\_$3, with the complex V$\_$2 moving successive-cyclically to
intermediate phase edge positions, until it finally merges with the
CP. Then, the second, intermediate, CP is generated, with V$\_$1
excorporating from the complex {\itshape [$\_${V$\_$2} V$\_$1 V$\_$2~]} category and
moving via the intermediate CP's phases edges until, finally, the
intermediate CP has been completed and V$\_$1 can take this CP as its
internal argument, via reprojection. The Tsakhur example in \ref{tsa1}
and the Hinuq example in \ref{tsa2} 
fully correspond to this scenario, with V$\_$1, V$\_$2, and V$\_$3 all
participating in agreement with DP$\_${abs}. However, examples
involving super-long-distance agreement like the one in 
\ref{tsa3} (from Hinuq) can also be found. 

\ea\label{ex:mueller:36} 
\gll Di\v{z} \label{tsa3}y-eq$'$i-yo [CP \textglotstop umar-i [CP Madina  y-aq$'$-e-s=t\l en~] ese-s-\l i~]\\
I.{\scshape dat} II-know-{\scshape prs} {} Umar-{\scshape erg} {} Madina(II) II-come-{\scshape pst}={\scshape quot} tell-{\scshape res}-{\scshape abst}\\
\glt `I know that Omar said that Madina came.'
\z
In (\Last), the intermediate verb V$\_$2 does in fact not exhibit overt
agreement marking even though both the matrix verb V$\_$1 and the most
deeply embedded verb V$\_$3 do. Still, (\Last) does not call into question
the present approach: It can plausibly be assumed that $\phi$-feature
agreement is indeed present on V$\_$2, but fails to be registered overtly
(there are many verbs that fail to exhibit visible agreement marking
despite showing up in the proper syntactic context in
Nakh-Daghestanian languages, and the reason for this is presumably
simply a morphological one). Thus, all in all, super-long-distance
agreement can be derived.\footnote{It should be mentioned that there
  are two further complications, though. First, recall that the
  present account of the obligatory information-structural reflex of
  long-distance agreement in terms of the Activity Condition would,
  strictly speaking, require two different additional features (next
  to the $\phi$-probes) on V$\_$1 and V$\_$2, and not just one, as in the
  cases discussed so far. It is not a priori clear what
  this extra feature might be. However, it has been argued that
  information-structural features like topic and focus do not qualify
  as primitives, but are rather composed of more primitive binary
  features (so as to capture natural classes of information-structural
  categories), like [$\pm$new], [$\pm$prom] (with, say, topic emerging
  as [--new,+prom]); see \cite{Choi:99}, based on
  \cite{Vallduvi:92}). If so, V$\_$2 and V$\_$1 can be equipped with
  separate pieces of [inf-st] information.

Second, \cite{Forker:11} also maintains that it is not completely
impossible in Hinuq to have super-long-distance agreement involving
V$\_$1, V$\_$3, and DP$\_${abs} in the most deeply embedded clause, not merely
in the absence of agreement on V$\_$2 (as in \ref{tsa3}), but in the presence of a {\it
  different} agreement on V$\_$2.  If such sentences (which  Forker assigns
an intermediate status, signalled by ``?'') can be substantiated as
grammatically well formed, additional assumptions that complement the
present analysis will be called for.}

A further property of long-distance agreement that needs to be
accounted for concerns \citegen{Polinsky&Potsdam:01} observation
that the C element {\itshape \textcrlambda in} blocks the operation in Tsez
(cf. section 2.4 above). Given that there is good evidence that
long-distance agreement across CP is possible in principle in
Nakh-Daghestanian languages, and given that we have analyzed the
transparent morpheme {\itshape {\l}i} as a C item, too, a recourse to a
general blocking nature of C is not available in the present
approach. Also, it is not possible to claim that a reprojecting V$\_$1
cannot merge with a CP headed by {\itshape \textcrlambda in}: First, the
[{\small $\bullet$}C{\small $\bullet$}] feature responsible for
reprojection movement is not sensitive to a difference between C
heads, and an additional selection relation (mediated by Agree) would
have to be stipulated; second (and more importantly), [{\small
    $\bullet$}C{\small $\bullet$}] on a reprojecting V is exactly the
same feature as [{\small $\bullet$}C{\small $\bullet$}] on a V that
fails to undergo complex predicate formation, and successfully takes
CP complements headed by {\itshape \textcrlambda in} in environments
without long-distance agreement. In view of this, we would like to
suggest that the blocking effect of a C head {\itshape \textcrlambda in} is
due to the fact that it does not permit a specifier. Thus, the problem
with long-distance agreement in these contexts can be traced back to
the unavailability of intermediate movement of a V$\_$1 that is initially
part of a complex predicate, to SpecC: As a consequence, the final
reprojection step of V$\_$1 will have to fatally violate the PIC (V$\_$1
can only reach SpecT, which is not accessible anymore once CP has been
completed).

Finally, we would like to point out that the present approach in terms
of reprojection makes a very simple prediction: Reprojection movement
of a verb by definition creates a head-complement structure; there is
no way how a specifier or adjunct could be involved (since this would
require a non-X$^0$ category to move). Therefore, long-distance
agreement is expected never to occur into subject clauses or adjunct
clauses.  This prediction is borne out: Long-distance agreement always
involves complement clauses.


\section{Conclusion}

We have argued that from the point of view of a model of syntax where
all operations apply in strictly local domains (as defined by the
Phase Impenetrability Condition (PIC)), and in the face of empirical
evidence showing that long-distance agreement can involve a matrix
verb and an agreement-controlling DP separated by a CP, none of the
existing approaches to long-distance agreement (non-local analyses,
small structure analyses, cyclic Agree analyses, and analyses where
movement to the edge feeds agreement) work satisfactorily. In view of
this, we have developed a new approach in terms of pre-syntactic
complex predicate formation and reprojection: The derivation starts
out with a complex verb {\itshape V$\_$1-V$\_$2} headed by V$\_$2, so that agreement
of V$\_$1 with DP can apply early in the derivation (not late, as in
other approaches), in an extremely local domain, and subsequent
reprojection movement of V$\_$1 turns the latter into a matrix verb, thereby
masking the locality of agreement and creating opacity (viz.,
counter-bleeding) in syntax.  

This approach may at first sight look quite radical. However, it is
worth bearing in mind that it suggests itself without further ado once
two widely employed operations are adopted and combined, viz., (i)
pre-syntactic complex predicate formation, and (ii) head movement as
reprojection. The properties that these two operations must have for
the analysis to work all qualify as independently motivated, and they
often correspond to standard assumptions in the field (in analyses
that adopt the operations). As a matter of fact, the only innovative
assumption that we have come up with is that pre-syntactic complex
predicate formation does not (or does not have to) result in a single
list of structure-building and agreement-inducing features (via a
process of functional composition), but can maintain the integrity and
independence of the two individual lists of structure-building and
agreement-inducing features.

Nevertheless, ideally there should be independent evidence for the
type of interaction of complex predicate formation and reprojection
movement that is at the heart of the present analysis of long-distance
agreement. To end this paper, we would like to briefly sketch an
approach to an entirely different phenomenon that works in the same
way, viz., extraction from DPs in German. 

As for the empirical evidence, extraction is impossible from subject
DPs and indirect object (dative-marked) DPs. This is shown (with
wh-movement as the extraction operation) in (\Next[a]) and (\Next[b]),
respectively.

\ea\label{ex:mueller:37}
\ea[*]{
\gll\relax [PP \"Uber wen~]$\_$1 \label{35a}hat [DP ein Buch t$\_$1~] den Karl beeindruckt~?\\
{} about whom has {} a book-{\scshape nom} {} the Karl-{\scshape acc} impressed\\
\glt `For which person is it the case that a book about that person impressed Karl?'}
\ex[*]{[PP \"Uber wen~] hat \label{35b}sie [DP einem Buch t$\_$1~] keine Chance gegeben~?\\
{} about whom has she-{\scshape nom} {} a book$\_${dat} {} no chance-{\scshape acc} given\\
\glt `For which person is it the case that she gave a book about that person no chance?'}
\z 
\z
With extraction from direct object (accusative-marked) DPs, things are
somewhat more variable: With some combinations of V and N, extraction
is possible (see \Next[a]), with other combinations, it is not (see
\Next[b]). 

\ea\label{ex:mueller:38}
\ea[]{
\gll\relax [PP \"Uber wen~]$\_$1 \label{36a}hat Karl [DP ein Buch t$\_$1~] gelesen~?\\
{} about whom has Karl-{\scshape nom} {} a book-{\scshape acc} {} read\\
\glt `Who did Karl read a book about?'}
\ex[*]{
\gll\relax [PP \"Uber wen~]$\_$1 \label{36b}hat Karl [DP ein Buch t$\_$1~] geklaut~?\\
{} about whom has Karl-{\scshape nom} {} a book-{\scshape acc} {} stolen\\
\glt `For which person is it the case that Karl stole a book about that person?'}
\z
\z
Thus, both structural and lexical factors play a role: On the one
hand, extraction from DP can be well formed in German if DP is a complement
(as with direct objects in \Last[ab]), but not if it is a specifier (as in
with subjects and indirect objects in (\LLast[a--b]), which can be assumed
to occupy Specv and SpecAppl positions, respectively). On the other
hand, extraction from DP also requires V and N to form a tight unit,
or a ``natural predicate''. This latter status is arguably determined
both by semantic considerations and by extralinguistic factors
(frequency, entrenchment), and it may to some extent vary from speaker
to speaker. Still, it must be modelled in the grammar in some way. In
\cite{Mueller:91:abs} and \cite{Mueller&Sternefeld:95}, it is proposed
that the relevant concept is that of abstract incorporation (in
\citegen{Baker:88} sense, conceived of as incorporation at LF that
is signalled already by co-indexation of heads in overt syntax): V
({\itshape read}) and N ({\itshape book}) in (\Last[a]) undergo abstract
incorporation and thus form a natural predicate, whereas V ({\it
  steal}) and N ({\itshape book}) do not (for most speakers). Given that
the theory of locality constraints on movement is sensitive to this
difference (as well as to the structural difference between
complements and specifiers), the data in (\LLast) and (\Last) can then all
be accounted for. Similar approaches in terms of abstract
incorporation have subsequently been developed by
\cite{Davies&Dubinsky:03} and \cite{Schmellentin:06}. However, there
are problems with this kind of approach. In particular, on this view
abstract incorporation of N into V must either be able to apply
non-locally, across an intervening DP projection (plus, possibly,
other functional projections in the DP that may intervene between D
and N); or the analysis must abandon the DP-over-NP hypothesis. To be
sure, there are ways out for the abstract incorporation
approach.\footnote{For instance, in \cite{Mueller:11:loc}, abstract incorporation is
  viewed as a regular syntactic Agree operation, with no actual
  movement involved.} Still, it can be noted that an approach based on
complex predicate formation plus reprojection can account for the data
in a very simple way. 

Thus, suppose that some combinations of V and N can undergo
pre-syntactic (lexical) complex predicate formation whereas others
cannot do so. This means that complex heads like the one in (\Next) can
be primitive inputs of Merge operations in the syntax.\footnote{As
  before, the head that will ultimately come to occupy a higher
  (c-commanding) position must be the one that fails to project in a
  well-formed derivation; cf. footnote \ref{vari2}.}

\ea\relax\label{ex:mueller:39} [N$\_$2 V$\_$1 N$\_$2~] \z
In the ensuing derivation, N$\_$2 first discharges its structure-building
and probe features (thereby undergoing Merge with a PP); see (\Next[a]).
Then
DP is added on top of NP (and possibly also other functional projections
before that), with V$\_$1 undergoing intermediate, Last Resort-driven
movement to SpecD; see (\Next[a]). After that, V$\_$1 undergoes the
final  movement step to take DP as its complement and thereby
discharge its [{\small $\bullet$}D{\small $\bullet$}] feature; see
(\Next[b]). From this point onwards, everything proceeds exactly as in
a derivation where there is no complex predicate formation as in
(\Next); such a derivation produces the VP in (\NNext). 

\ea\label{ex:mueller:40}
\ea\relax [NP [N$\_$2 V$\_$1 N$\_$2~] PP~]
\ex\relax [DP V$\_$1 [D$'$ D [NP [N$\_$2 t$\_$1 N$\_$2~] PP~]]]
\ex\relax [VP V$\_$1 [DP t$'$\_$1$ [D$'$ D [NP [N$\_$2 t$\_$1 N$\_$2~] PP~]]]]
\z
\z

\ea\label{ex:mueller:41} [VP V$\_$1 [DP D [NP N$\_$2 PP~]]]\z
Importantly, both the NP and the PP in (\LLast[c]) (based on complex
predicate formation of V and N) have been in an extremely local
(object-like) relation with V whereas the NP and PP in (\Last) (based
on regular, separate projection of V and N) have never been in a local
relation with V. Without going into the details of how exactly this
will best be implemented in a given theory of locality restrictions on movement, it
seems plausible to assume that it is the extremely local relation of V
and NP/PP at an earlier derivational step that makes extraction of PP
from DP possible in  \ref{36a}, and it is the absence of such a relation that blocks
the movement in \ref{36b}. As before, reprojection movement of V
by definition cannot take place from specifier, which then accounts
for the illformedness of \ref{35a} and \ref{35b}.

\section*{Abbreviations}
\begin{multicols}{2}
\begin{tabbing}
\textsc{intrans}\hspace{5mm} \= information structure\kill
\textsc{abs} \> absolutive\\ 
\textsc{abst} \> abstract suffix\\ 
\textsc{acc} \> accusative\\
\textsc{caus} \> causative\\ 
\textsc{cl} \> clitic\\ 
\textsc{cop} \> copula\\ 
\textsc{cvb} \> narrative converb\\ 
\textsc{dat} \> dative\\ 
\textsc{erg} \> ergative\\ 
\textsc{fem} \> feminine\\ 
\textsc{foc} \> focus\\ 
\textsc{inf} \> infinitive\\ 
\textsc{inf-st} \> information structure\\ 
\textsc{intrans} \> intransitive\\ 
\textsc{lat} \> lative\\ 
\textsc{masc} \> masculine\\ 
\textsc{nom} \> nominative\\ 
\textsc{nmlz} \> nominalizer\\ 
\textsc{obj} \> object\\ 
\textsc{obl} \> oblique\\
\textsc{perf} \> perfective\\
\textsc{pl} \> plural\\ 
\textsc{pr(e)s} \> present\\ 
\textsc{pst} \> past\\ 
\textsc{pstprt} \> past participle\\
\textsc{ptcp} \> participle\\ 
\textsc{q} \> question particle\\ 
\textsc{quot} \> quotative enclitic\\
\textsc{res} \> resultative participle\\ 
\textsc{sg} \> singular\\ 
\textsc{subj} \> subject\\ 
\textsc{ti} \> transitive inanimate
\end{tabbing} 
\end{multicols}

\section*{Acknowledgments}
For comments and discussion, we are grateful to Josef Bayer, Elena
Pyatigorskaya, Tonjes Veenstra, Martina Martinovi\'c, Johannes Mursell
(and the other editors of the present volume), and audiences at the
{\itshape GGS} meeting at Universit{\"a}t Konstanz (2014), the workshop on
{\itshape Agreement} at the University of York (2014), and the workshop
{\itshape Heads, Phrases, Ts and Nodes} at Humboldt Universit{\"a}t Berlin (2017).

{\sloppy
\printbibliography[heading=subbibliography,notkeyword=this]}
\end{document}
